---
title: "semhelpinghands"
author: "Shu Fai Cheung"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{semhelpinghands}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

At the time of writing,
[`semhelpinghands`](https://sfcheung.github.io/semhelpinghands/)
has these
groups of functions: manipulating
parameter estimates tables, comparing
results from different methods,
bootstrapping, and ... others.

This article will only introduce very
briefly the groups. Some of them will be
introduced in details in forthcoming
article dedicated to them.

Let's load the package first, and
also load `lavaan`.

```{r setup}
library(semhelpinghands)
library(lavaan)
```

# Manipulate Parameter Estimates Tables

In using `lavaan`, I prefer reading the
output of `parameterEstimates()`, which
is compact and clear to me. I sometimes
would like to organize the rows and
columns in ways meaningful to a
particular research questions. This
can certainly be done using base R
or `dplyr`. However, I am lazy and want
to be able to do things using just one
or two functions, with just one or two
arguments.

This is a sample dataset for illustration,
`dvs_ivs`,
with 3 predictors (`x1`, `x2`, and `x3`),
3 outcome variables (`y1` `y2`, and `y3`),
and a group variable (`gp`).

First a single sample model:

```{r}
data(dvs_ivs)
mod <-
"
y1 ~ x1 + x2 + x3
y2 ~ x1 + x3
y3 ~ y2 + x2
"
fit <- sem(model = mod,
           data = dvs_ivs,
           fixed.x = FALSE)
```

The parameter estimates tables:

```{r}
est <- parameterEstimates(fit)
est
```

A two-sample model is also fitted to the
dataset:

```{r}
fit_gp <- sem(model = mod,
              data = dvs_ivs,
              group = "gp",
              fixed.x = FALSE)
```

This is the parameter estimates table:

```{r}
est_gp <- parameterEstimates(fit_gp)
est_gp
```

So, these are what `semhelpinghands`
has for now:

## Add Significance Test Results: `add_sig()`

Despite the controversy over null
hypothesis significance testing,
we still need to report them, for now.
They are there, but I have to decode
them mentally. Here comes `add_sig()`:

```{r}
add_sig(est)
```

If bootstrapping was used, it also
supports using the confidence intervals,
which may yield results different from
the *p*-values when bootstrapping
confidence intervals are used
(but same results in this example):

```{r}
add_sig(est,
        use = c("pvalue", "ci"))
```

It also works on the standardized
solution:

```{r}
std <- standardizedSolution(fit)
add_sig(std)
```

Note: But be careful about the
interpretation of the *p*-values
in the standardized solution,
which are based on the delta method.
See `vignette("standardizedSolution_boot_ci")`.

See the help page of `add_sig()`
for other options.


## Filter by Operators and Some Other Columns: `filter_by()`

Its purpose is very simple, selecting
rows by commonly used columns: operators (`op`),
"dependent variables" (`lhs`), and
"independent variables" (`rhs`).

```{r}
filter_by(est,
          op = "~")
```

It also supports filtering by group
using group labels:

```{r}
filter_by(est_gp,
          op = "~",
          group = "gp1",
          fit = fit_gp)
```

See the help page of `filter_by()`
for other options.


## Group By DVs or Group By IVs: `group_by_dvs()` and `group_by_ivs()`

Sometimes I want the conventional
iv-column-dv-row or dv-column-iv-row
format. That's what `group_by_dvs()`
and `group_by_ivs()` are for:

```{r}
group_by_dvs(est)
group_by_ivs(est)
```

They also supports extracting another
column:

```{r}
group_by_dvs(est,
             col_name = "pvalue")
group_by_ivs(est,
             col_name = "pvalue")
```

See the help page of `group_by_dvs()`
and `group_by_ivs()`
for other options.


## Group By Groups: `group_by_groups()`

In multiple sample models, one common
task is comparing results across groups.
I wrote `group_by_groups()` for this
task, to compare results side-by-side:

```{r}
group_by_groups(est_gp)
```

It also supports extracting several
columns:

```{r}
group_by_groups(est_gp,
                col_names = c("est", "pvalue"))
```

If the fit object is used, it can print
group labels:

```{r}
group_by_groups(fit_gp,
                col_names = c("est", "pvalue"))
```

See the help page of `group_by_groups()`
for other options.


## Compare Models: `group_by_models()`

This is inspired by the proposal
Rönkkö proposed in a GitHub
[issue](https://github.com/simsem/semTools/issues/24#issue-235172313)
for `semTools`. I want something
simple for a quick overview and so
I wrote `group_by_models()`.

Suppose this is the other model fitted:

```{r}
mod2 <-
"
y1 ~ x1 + x2 + x3
y2 ~ x1 + x2
y3 ~ y2 + x1
"
fit2 <- sem(model = mod2,
            data = dvs_ivs,
            fixed.x = FALSE)
est2 <- parameterEstimates(fit2)
est2
```

These two models have no nested
relationships. To compare the estimates,
`group_by_models()` can be used:

```{r}
group_by_models(list(Model1 = est,
                     Model2 = est2))
```

It can also compare several columns:

```{r}
group_by_models(list(Model1 = est,
                     Model2 = est2),
                col_names = c("est", "pvalue"))
```

See the help page of `group_by_models()`
for other options.

## Sort Rows: `sort_by()`

The function `sort_by()` can be used
to sort rows using (a) common fields
such as `"op"`, `"lhs"`, and `"rhs"`,
(b) operators such as `"~"` and `"~~"`.
It can be used on the output of
some other functions that manipulate
a parameter estimates table.

```{r}
out <- group_by_groups(est_gp,
                       col_names = c("est", "pvalue"))
out <- filter_by(out,
                 op = c("~", "~~"))
sort_by(out,
        by = c("op", "rhs"))
```

Some functions, such as `group_by_models()`,
will automatically
call `sort_by()` to sort the results.

The default order should be acceptable
in most cases, but can also be customized.
See the help page of `sort_by()`
on customizing the order.

## Piping

Though not officially supported, piping using
`|>` can be used with most of the
functions that manipulate
parameter estimates tables. For example:

```{r}
est_gp |>
  add_sig() |>
  group_by_groups(col_names = c("est", "pvalue", "sig"),
                  group_first = FALSE) |>
  filter_by(op = c("~"))
```


# Compare Methods

Though I believe the choice of the
estimation method should be justified,
suppose we want to assess the sensitivity
of the parameter estimate results to
the methods used, `compare_estimators()`
can be used as a quick way to compare
results from different methods.

```{r}
out <- compare_estimators(fit,
         estimator = c("ML", "GLS", "MLR"))
group_by_models(out,
                col_names = c("se", "pvalue"))
```

It simply refits the models for each
estimator and returns the results.
They can then be treated as different
"models" and processed by `group_by_models()`.

`se_ratios()` is a wrapper of `group_by_models()`
used to compare the standard errors
by different estimators in the output
of `compare_estimators()`:

```{r}
se_ratios(out,
          reference = "ML")
```

See the help page of `compare_estimators()`
for other options.

# Bootstrapping

One issue with the standardized solution
is the confidence intervals. They are
based on the delta method even if
`se = "boot"` is used. For indirect
effects, for which bootstrap confidence
intervals are commonly used, the
confidence intervals for them in
the standardized solution are not
what usually reported other tools
for mediation. There are some
other powerful tools on the Internet and
the [Google Group for lavaan], see
[this thread](https://groups.google.com/g/lavaan/c/0RSsh4M6zQg/m/I85YFAiLAAAJ),
to address this problem.
I wrote `standardizedSolution_boot_ci()`
not to replace them (it obviously can't),
but to address a very specific
case I usually encounter myself:
Generating the bootstrap confidence
intervals for standardized estimates
based on the bootstrap estimates already
generated by `se = "boot"`.

Please see `vignette("standardizedSolution_boot_ci")`
for an illustration on `standardizedSolution_boot_ci()`.

# Others

## Showing the Options in a Model

`lavaan::lavaan()` and its wrappers suc has `lavaan::sem()`
and `lavaan::cfa()` allow users to set several options
using an `estimator`: `ML`, `GLS`, `WLSMV`, and others.
However, it is not easy to remember
what the options set for an estimator.
Instead of finding them from the output
of `summary()`, this function shows some
of them in one table for a quick overview.
This is an example:

```{r}
data(dvs_ivs)
mod <-
"
y1 ~ x1 + x2 + x3
y2 ~ x1 + x3
y3 ~ y2 + x2
"
fit_default <- sem(model = mod,
                   data = dvs_ivs)
show_more_options(fit_default)
fit_MLR <- sem(model = mod,
               data = dvs_ivs,
               estimator = "MLR")
show_more_options(fit_MLR)
fit_MLR_fiml <- sem(model = mod,
                    data = dvs_ivs,
                    estimator = "MLR",
                    missing = "fiml")
show_more_options(fit_MLR_fiml)
```

## Recoding Minimization History

In structural equation modeling,
closed-form solution is rare and
optimization (minimization) is used
to find the/a solution. Out of curiosity
and teaching, I wrote a function
to capture the minimization history
so I can examine it or even plot
the process. The function,
`record_history()`, is still in early
development but should work for now
for common simple scenarios.
Please refer the help page of
`record_history()`
to learn more about it.

## Add Covariances Between "Exogenous" Variables

In `lavaan`, I rarely need to manually add
covariances between exogenous variables
(defined in a loose sense: variables appear
on the right-hand side but not on the left-hand
side of `~`). However, I came into a situation
in which `lavaan` will not do this
(for good reasons). For example,
when a covariance between a residual
term and an exogenous variable is set
to free. I wrote two simple functions,
`add_exo_cov()` and `auto_exo_cov()`,
for this purpose. Please refer to
their help pages for further information.

